---
title: "Ch11"
author: "George Athanasopoulos"
date: "15/01/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fpp3)
library(tictoc)
```

# Intro 

- We will learn/review hierarchical and grouped structures
- The concept of coherence
- Learn some useful commands/functions within the fable package 

## Hierarchical and grouped time series {#hts}

### Hierarchical time series {-}

### Example: Australian tourism hierarchy {-}

```{r recode}
tourism <- tsibble::tourism %>%
  mutate(State = recode(State,
    `New South Wales` = "NSW",
    `Northern Territory` = "NT",
    `Queensland` = "QLD",
    `South Australia` = "SA",
    `Tasmania` = "TAS",
    `Victoria` = "VIC",
    `Western Australia` = "WA"
  ))

tourism
# is a tsibble with 304 series 
# with Key variables Region, State, Purpose
tourism$Quarter %>% range()
```

- The first important function `aggregate_key()`
- Use `parent/child` specification for hierarchical/nested time series.

```{r nested, echo=TRUE}
tourism_hts <- tourism %>%
  aggregate_key(State / Region, Trips = sum(Trips))

tourism_hts
# additional rows corresponding to state and national aggregations for each quarter.
# note <chr*> signal an aggregation structure
```

- The second important function `is_aggregated()` which allows you to filter through aggregation structures

```{r tourismStates, fig.width=9, fig.asp=0.7, fig.cap="Domestic overnight trips from 1998Q1 to 2017Q4 aggregated by state.", warning=FALSE, message=FALSE, echo=TRUE}
tourism_hts %>%
  filter(is_aggregated(Region)) %>% # select everything but not Region
  autoplot(Trips) +
  labs(y = "Trips ('000)",
       title = "Australian tourism: national total and states") +
  facet_wrap(vars(State), scales = "free_y", ncol = 3) +
  theme(legend.position = "none")
```

Let's have another look using `filter()` - showing some important dynamics.

```{r seasonStates, echo=FALSE, fig.cap="Seasonal plots for overnight trips for the Northern Territory and Queensland, and Tasmania and Victoria highlighting the contrast in seasonal patterns between northern and southern states in Australia.", fig.asp=0.5, fig.width=7, out.width="80%", message=FALSE, warning=FALSE}
tourism_hts %>%
  filter(State == "NT" | State == "TAS" |
           State == "QLD" | State == "VIC", is_aggregated(Region)) %>%
  select(-Region) %>% # if you don't do this R will complain
  gg_season(Trips) +
  facet_wrap(vars(as.character(State)), ncol = 2, scales = "free_y")+
  labs(y = "Trips ('000)")
```
These show the importance of modelling all series individually as signals will be lost at the aggregate level or too noisy at levels below as we see in Figure 11.5.

### Grouped time series {-}

### Example: Australian prison population {-}

- Using `aggregate_key()` for grouped/crossed structure
- `attribute1*attribute2`

```{r prisongts, fig.width=9, fig.asp = .7, echo=FALSE, fig.cap="Total Australian quarterly adult prison population, disaggregated by state, by legal status, and by gender.", warning=FALSE, message=FALSE, fig.pos="b", fig.env="figure*"}

prison <- readr::read_csv("https://OTexts.com/fpp3/extrafiles/prison_population.csv") %>%
  mutate(Quarter = yearquarter(Date)) %>%
  select(-Date) %>%
  as_tsibble(key = c(Gender, Legal, State, Indigenous), index = Quarter) %>%
  relocate(Quarter) # just relocating Quarter as the first column

prison
# 64 series 8*2*2*2

prison_gts <- prison %>%
  aggregate_key(Gender * Legal * State, Count = sum(Count) / 1e3)


prison_gts
# now 81 series 
# Aggregate 1 series
# 1-way 8+2+2 series
# 2-way 8*2+8*2+2*2=36
# Bottom-level 8*2*2=32 series
# Total: 81=1+(8+2+2)+(8*2+2*2+2*8)+32
```

We'll look more carefully at these when we do the forecasting.

### Mixed hierarchical and grouped structure {-}

- nested and crossed disaggregating factors

```{r mixed, echo=TRUE}
tourism_full <- tourism %>%
  aggregate_key((State / Region) * Purpose, Trips = sum(Trips))

tourism_full 
# 425 series
# 85 from hierarchical
# 340 = 85*4 from crossings
```

Lets have a look at the purpose of travel series 

```{r mixed-purpose, fig.width=10, fig.asp = 0.6, echo=FALSE, fig.cap="Australian domestic overnight trips from 1998Q1 to 2017Q4 disaggregated by purpose of travel.", message=FALSE, warning=FALSE, dependson="mixed",fig.env="figure*"}
tourism_full %>%
  filter(is_aggregated(State), 
         is_aggregated(Region), 
         !is_aggregated(Purpose)) %>%
  ggplot(aes(x = Quarter, y = Trips,
             group = as.character(Purpose), colour = as.character(Purpose))
  ) +
  stat_summary(fun = sum, geom = "line") +
  facet_wrap(~ as.character(Purpose), scales = "free_y", nrow = 2) +
  labs(title = "Australian tourism: by purpose of travel",
       y = "Trips ('000)") +
  # theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  guides(colour = guide_legend("Purpose"))
```

# Generating coherent forecasts

- Third important function `reconcile()`
- Note the workflow 

```r
data %>% aggregate_key() %>% model() %>% reconcile() %>% forecast()
```

1. Begin with a `tsibble` object (here labelled `data`) containing the individual bottom-level series.
2. Define in `aggregate_key()` the aggregation structure and build a `tsibble` object that also contains the aggregate series.
3. Identify a `model()` for each series, at all levels of aggregation.
4. Specify in `reconcile()` how the coherent forecasts are to be generated from the selected models.
5. Use the `forecast()` function to generate forecasts for the whole aggregation structure.


## Single level approaches {#single-level}

### The bottom-up approach {-}

### Example: generating bottom-up forecasts

Let's start with a simple example

```{r tourism_states, message=FALSE}
tourism_states <- tourism %>%
  aggregate_key(State, Trips = sum(Trips))
```

```{r bottom_up, message=FALSE, cache=TRUE}
fcasts <- tourism_states %>%
  model(ets = ETS(Trips)) %>%
  reconcile(bu = bottom_up(ets)) %>%
  forecast()

fcasts
# This may be a little tricky at first
# 18 series comprise
# 2 sets of forecasts (ets + bu) * (8 states + 1 aggregate)

fcasts %>% filter(is_aggregated(State)) # Aggregate
fcasts %>% filter(!is_aggregated(State)) # Bottom level
```

```{r top_down, message=FALSE, cache=TRUE}
fcasts <- tourism_states %>%
  model(ets = ETS(Trips)) %>%
  reconcile(bu = bottom_up(ets),
            av_prop = top_down(ets, method="average_proportions"),
            prop_av = top_down(ets, method="proportion_averages"),
            # fcast_prop = top_down(ets, method="forecast_proportions") 
            # not yet supported
            ) %>%
  forecast()

fcasts$.model %>% unique()
```

## Forecast reconciliation {#reconciliation}

### The MinT optimal reconciliation approach {-}

MinT (Minimum Trace) optimal reconciliation approach is implemented by `min_trace()` within the `reconcile()` function.

1. OLS: `method = "ols"`
2. WLS(variance scaling) `method = "wls_var"`
3. WLS(structural scaling) `method = "wls_struct"`
4. MinT(Sample) `method = "mint_cov"`
5. MinT(Shrink) `method = "mint_shrink"`

## Forecasting Australian domestic tourism {#tourism}

```{r tourismfit, echo=TRUE}
tourism_full <- tourism %>%
  aggregate_key((State / Region) * Purpose, Trips = sum(Trips))

#tic()
# DO NOT RUN (187.7 sec elapsed)
fit_tourism <- tourism_full %>%
  filter(year(Quarter) <= 2015) %>%
  model(base = ETS(Trips)) %>%
  reconcile(
    bu = bottom_up(base),
    ols = min_trace(base, method = "ols"),
    mint = min_trace(base, method = "mint_shrink"),
  )
#toc()
fit_tourism
fit_tourism %>% filter(Region=="Canberra")

#tic()
# DO NOT RUN (takes 46.51 sec elapsed)
fc_tourism <- fit_tourism %>% forecast(h = "2 years")
#toc()

fc_tourism
fc_tourism %>% filter(is_aggregated(Region), is_aggregated(Purpose))
fc_tourism$.model %>% unique()
fc_tourism %>% distinct(State, Purpose) #5*9 
```

```{r tourism-states, fig.width=10, fig.asp = .6, echo=TRUE, fig.cap="Forecasts of overnight trips for Australia and its states over the test period 2016Q1--2017Q4.", warning=FALSE, message=FALSE, fig.pos="b", fig.env="figure*", dependson="tourismfc"}
fc_tourism %>%
  filter(is_aggregated(Region), is_aggregated(Purpose)) %>%
  autoplot(
    tourism_full %>% filter(year(Quarter) >= 2011),
    level = NULL
  ) +
  labs(y = "Trips ('000)") +
  facet_wrap(vars(State), scales = "free_y")
```

```{r tourism-purpose, echo=FALSE, fig.asp=.5, echo=TRUE, fig.cap="Forecasts of overnight trips by purpose of travel over the test period 2016Q1--2017Q4.", fig.env="figure*", fig.pos="b", fig.width=10, message=FALSE, warning=FALSE, dependson="tourismfc"}
fc_tourism %>%
  filter(is_aggregated(State), !is_aggregated(Purpose)) %>%
  autoplot(
    tourism_full %>% filter(year(Quarter) >= 2011),
    level = NULL
  ) +
  labs(y = "Trips ('000)") +
  facet_wrap(vars(Purpose), scales = "free_y")
```

Let's do some evaluation

```{r fcaccuracy2, message=FALSE, dependson="tourismfc"}
fc_tourism %>%
  filter(is_aggregated(State), is_aggregated(Purpose)) %>%
  accuracy(
    data = tourism_full,
    measures = list(rmse = RMSE, mase = MASE)
  ) %>%
  group_by(.model) %>%
  summarise(rmse = mean(rmse), mase = mean(mase))
```

Go to book and see complete forecast evaluation Table 11.2.

## Reconciled distributional forecasts {#rec-prob}

## Forecasting Australian prison population {#prison}

```{r prisonfc}
#
# DO NOT RUN
#
fit_prison <- prison_gts %>%
  filter(year(Quarter) <= 2014) %>%
  model(base = ETS(Count)) %>%
  reconcile(
    bottom_up = bottom_up(base),
    MinT = min_trace(base, method = "mint_shrink")
  )
fc_prison <- fit_prison %>% forecast(h = 8)

fit_prison
fc_prison #1944 rows = 81*8*3
```

Revert to the book for the analysis of the forecasts/plots. 

