---
title: "Bookclub"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simple linear regression

```{r}

# Data prepared

library(readxl)
LungCapData <- read_excel("~/Desktop/bookclub/LungCapData.xls")
View(LungCapData)
attach(LungCapData)
LungCapData$Smoke <- as.factor(Smoke)
names(LungCapData)

plot(Age, LungCap, main = "The relationship between lung capacity and age")

# Pearson's correlation 

cor(Age, LungCap)

# Fitting linear models

mod <- lm(LungCap ~ Age)

summary(mod)

# extract the coefficients
attributes(mod)

mod$coefficients

#  plot the model

plot(Age, LungCap,pch = 21, main = "The relationship between lung capacity and age")
abline(mod, col = 'red', lwd = 3)

# confidence interval

confint(mod)

confint(mod,level = 0.99)

summary(mod)

# ANOVA table
anova(mod)
```
## Checking linear Regression Assumptions in R

```{r}
mod <- lm(LungCap ~ Age)

plot(mod)

## non-linear regression plot

set.seed(10)
y = rnorm(50, 0, 200)
x = runif(50, 0,100)
plot(x,y)

nolinear <- lm(y~x)


plot(y,x)
abline(nolinear)


plot(nolinear)
```

## Multiple Linear Regression in R

$\mathrm{y}(\mathrm{x})=\beta_{0}+\beta_{1} \mathrm{x}_{1}+\beta_{2} \mathrm{x}_{2}+\beta_{3} \mathrm{x}_{3}+\ldots . \beta_{\mathrm{n}} \mathrm{X}_{\mathrm{n}}$

```{r}
# fit a model using age and height as X-variables
model1 <- lm(LungCap ~ Age + Height)

# get a summary of the model
summary(model1)

# calculate Pearson's correlation between Age and Height
cor(Age, Height,method= "pearson")

# ask for confidence intervals for the model coefficients
confint(model1, conf.level = 0.95)

# fit a model using all X variables
model2 <- lm(LungCap ~ Age + Height+ Smoke + Gender +Caesarean)

# summary of the model
summary(model2)

# check the regression diagnostic plots for this model
plot(model2)


```




$$
\hat{\mu}_{y \mid x}=1.08+0.555 \times \text { Age }-0.649 \times \text { Smoke }
$$

## Dummy variable

```{r}

class(Smoke)
levels(Smoke)

# build the regression model
Smoke <- relevel(Smoke, ref = "yes")
table(Smoke)
mod1 <- lm(LungCap~ Age+ Smoke)
summary(mod1)

# including categorical variable into R
plot(Age[Smoke =="no"], LungCap[Smoke =="no"], col = "blue", ylim = c(0,15), xlab = "Age", ylab = "LungCap", main = "Lungcapacity VS Age, smoke")
points(Age[Smoke =="yes"],LungCap[Smoke =='yes'], col="red", pch=16)
legend(3,15, legend = c("NonSmoker","Smoker"), col = c("blue", "red"), pch = c(1,16), bty = "n")

# adding the regression line 
# regression line for the non-smokers, in blue.
abline(a=1.08, b=0.555, col = "blue",lwd = 3)
# Regression line for Smokers in red
abline(a = 0.413, b= 0.555, col = "red", lwd = 3)

# Change the reference group

class(Smoke)
Smoke <- as.factor(Smoke)
table(Smoke)
levels(Smoke)

mod <- lm(LungCap ~ Smoke)
summary(mod)

# set another reference
Smoke <- relevel(Smoke, ref = "yes")
Smoke <- relevel(Smoke, ref = "no")
table(Smoke)

mod1 <- lm(LungCap~ Age+ Smoke)
summary(mod1)

# transfer the numeric variable in to category variable
cHeight <- cut(Height, breaks = c(0,50,55,60,65,70,100), labels = c("A", "B", "C", "D", "E", "F"), right = FALSE)

Height[1:10]
cHeight[1:10]


#calculate the mean value  
mean(LungCap[cHeight = "A"])
mean(LungCap[cHeight = "B"])
mean(LungCap[cHeight = "C"])
mean(LungCap[cHeight = "D"])
mean(LungCap[cHeight = "E"])

# Generating the model
mod3 <- lm(LungCap ~ cHeight)

summary(mod3)


```
$$
X_{\text {smoke}}=\left\{\begin{array}{ll}
1, & \text { if they smoke } \\
0, & \text { otherwise }
\end{array}\right.
$$

$$
\begin{array}{l}
X_{B}=\left\{\begin{array}{ll}
1, & \text { if in Height Category } \mathrm{B} \\
0, & \text { otherwise }
\end{array}\right. \\
X_{C}=\left\{\begin{array}{ll}
1, & \text { if in Height Category } \mathrm{C} \\
0, & \text { otherwise }
\end{array}\right. \\
X_{D}=\left\{\begin{array}{ll}
1, & \text { if in Height Category } D \\
0, & \text { otherwise }
\end{array}\right. \\
X_{E}=\left\{\begin{array}{ll}
1, & \text { if in Height Category } \mathrm{E} \\
0, & \text { otherwise }
\end{array}\right. \\
X_{F}=\left\{\begin{array}{ll}
1, & \text { if in Height Category } F \\
0, & \text { otherwise }
\end{array}\right.
\end{array}
$$
An Individual in:
Height Category A will have: $X_{B}=0, X_{C}=0, X_{D}=0, X_{E}=0, X_{F}=0$
Height Category $\mathrm{B}$ will have: $X_{B}=1, X_{C}=0, X_{D}=0, X_{E}=0, X_{F}=0$
Height Category $\mathrm{C}$ will have: $X_{B}=0, X_{C}=1, X_{D}=0, X_{E}=0, X_{F}=0$
Height Category $\mathrm{D}$ will have: $X_{B}=0, X_{C}=0, X_{D}=1, X_{E}=0, X_{F}=0$
Height Category $\mathrm{E}$ will have: $X_{B}=0, X_{C}=0, X_{D}=0, X_{E}=1, X_{F}=0$
Height Category $\mathrm{F}$ will have: $X_{B}=0, X_{C}=0, X_{D}=0, X_{E}=0, X_{F}=1$

$\begin{aligned} \hat{\mu}_{L u n g C a p \mid x} &=b_{0}+b_{B} X_{B}+b_{C} X_{C}+b_{D} X_{D}+b_{E} X_{E}+b_{F} X_{F} \\ &=2.15+1.51 X_{B}+3.25 X_{C}+5.02 X_{D}+6.54 X_{E}+8.65 X_{F} \end{aligned}$

# select the best model

```{r}
# select the best model

# Example 1
# fit the full model
Full.model <- lm(LungCap ~ Age + Gender+ Smoke+ Height)
summary(Full.model)
# fit the reduced model
Reduced.model <- lm(LungCap ~ Age + Gender+ Smoke)
summary(Reduced.model)

# compare the two models by using the Partial F-test
anova(Reduced.model,Full.model)

``